---
title: "Calibration of Leicester score and enhanced scores"
author: "XL,TL,JC,LC,DH"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  word_document:
      fig_width: 8
      fig_height: 5
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

**Keywords**: 45% of the cohort, incident T2D derived from HES/Dth/PC censored till 2016.

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(kableExtra)
library(knitr)
library(yaml)
library(here)
library(ggplot2)
library(tidyr)
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(svglite)
library(rsvg)
library(png)
library(dplyr)
library(survival)
library(survminer)
library(Rcpp)
library(summarytools)
library(arsenal)
library(pander)
library(ggeffects)
library(sjPlot)
library(etm)
library(readr)
library(tibble)
library(ResourceSelection)
library(Publish)
library(rsample)
library(plotROC)
library(pROC)
library(nricens)
library(PredictABEL)
library(cutpointr)
library(caret)
library(gridExtra)
library(riskRegression)
library(gtools)
library(ggsurvfit)
# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
#options(knitr.kable.NA='', knitr.table.format = "html")
#options(kableExtra.auto_format = FALSE)
# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('missing', '')


knitr::opts_knit$set(root.dir = here::here())

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(config$functions))
source(here::here(config$scripts$nri))

my_controls <- tableby.control(
    test = F,
    total = T,
    #numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss"), 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      medianq1q3 = "Median (IQR)",
      range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )

# Read in data derived from code chunk below
data<-readRDS(here::here(file.path(config$data$derived,"data_Leicrs.RDS")))

tot_train<-readRDS(here::here(file.path(config$data$derived,"tot_train.rds")))
tot_test<-readRDS(here::here(file.path(config$data$derived,"tot_test.rds")))


```

```{r,include=FALSE}
source(file.path(config$scripts$cleaning, "dataset_generator.R"))

pretty_func <- pretty_switch(field_definitions=TEU_SPECS$T2DM_PRS, return_type = "function")
pretty_names <- pretty_switch(field_definitions=TEU_SPECS$T2DM_PRS, return_type="list")

pretty_names$hba1c_cat="HbA1c at baseline"
pretty_names$hba1c_cat2="HbA1c at baseline"

pretty_names$BSM_WaistCirc_Cat="Waist circumference,cm"
pretty_names$BSM_BMI_Cat="BMI"
pretty_names$HighBP="Antihypertensive meds or high blood pressure"
pretty_names$diab_rs_cat="Leicester risk score"
pretty_names$PRS_quintiles="T2D PRS quintiles"
pretty_names$Ethnicity="Ethnicity"

pretty_names$diab_rs="LRA Score"
pretty_names$Eth_rs="LRArev Score"
pretty_names$PRS_rs="LRAprs Score"
```


# Introduction

We are interested in assessing calibration of 10-year T2D risk. Since the maximum follow-up of quintile 2 of the enhanced score (with PRS) in test data was less than 10 years (see KM plot at the end of this document), we decided to compare predicted 10-year risk and observed 9-year risk (an approximate for observed 10-year risk). 


# Leicester score (Training)

First, we calculated the predicted probability from the intercept and beta coefficients of the Leicester score model. 

Then we followed the same approach for calibration discussed in [Sun2021](https://pubmed.ncbi.nlm.nih.gov/33444330/) where we compared the observed and predicted risks across deciles of predicted risk. 

```{r pred prob Leicrs}

tot_train<-tot_train%>%select(!starts_with("RS_"))

# Leicester risk score vars
columns=c("TEU_BaC_AgeCat","BaC_Sex","Ethnicity","TEU_FaH_diab","BSM_WaistCirc_Cat","BSM_BMI_Cat","HighBP")

rs_train<-cbind(tot_train, setNames(tot_train[columns], paste0("RS_", columns)))

rs_train<-rs_train%>%
  mutate_at(paste0("RS_",columns), ~var_to_beta(.))

# Sum all the columns starting with RS_

rs_train<-rs_train%>%
  mutate(sum_beta=-3.430668+rowSums(select(.,starts_with("RS_"))))%>%
  mutate(Leicrs_prob=exp(sum_beta)/(1+exp(sum_beta)))


```


```{r}
#Check if this probability actually makes sense 
tabs<-rs_train%>%group_by(diab_rs_cat)%>%summarise(p=mean(Leicrs_prob))

tabs<-tabs%>%mutate(
             p_website=case_when(diab_rs_cat=="Low:0-6" ~ 1/100,
                           diab_rs_cat=="Increased:7-15" ~ 1/35,
                           diab_rs_cat=="Moderate:16-24" ~ 1/10,
                           diab_rs_cat=="High:25-47" ~ 1/4))

pander(tabs,caption = "Predicted probability of the Leicester score model by risk groups. Column p shows the probability computed using the betas of Leicester score and column p_website shows the probability from the Diabetes UK website.")

#rs_train%>%group_by(diab_rs_cat)%>%summarise(n=n(),cases=sum(TEU_T2DM_status))%>%mutate(obs_p=cases/n)
```

(XL, 08Feb2023): Previously, we categorised participants into deciles groups then plotted the calibration but due to the small number of individuals with 10-year follow-up, we categorised participants into quintiles groups.

We first plotted the KM plot by quintiles of predicted probability from the Leicester model using the training data. 


```{r KM by quintiles}

rs_train<-rs_train%>%mutate(
    pred_groups=Percentile(Leicrs_prob,probs=seq(0, 1, 1/5),labels = c(1:5),right=TRUE))
  
survfit2(Surv(TEU_T2DM_time,TEU_T2DM_status) ~ pred_groups, data = rs_train) %>% 
  ggsurvfit() +
  labs(
    title = "KM plot by quintiles of predicted probability from the Leicester model \n using the training data",
    x = "Years",
    y = "Overall survival probability"
    ) + 
  add_confidence_interval() +
  scale_x_continuous(breaks = seq(0,12,2)) +
  add_risktable()

```


In the calibration plots below, each blue square represents a quintile group. The x-axis shows the mean of predicted probability (computed from betas of the Leicester model) within each quintile group and the y-axis shows the observed risk at 9 years estimated using UKB data within each quintile group.

Title: Calibration plot of the Leicester score with 10-year risk of T2D using the training data.

```{r Leicprob calibration}

Leic10_cal<-Calibration_df(data=rs_train,predRisk="Leicrs_prob",groups=5,times=9,format="wide")

#plot(pred_obs_wide$value.pred,pred_obs_wide$value.obs,main="",#xlab="",ylab="",
#pch=16,ps=2, xlim=c(0,0.4) , ylim=c(0,0.4),cex.lab=1.2,
#cex.axis=1.1, las=1)
#lines(x =c(0,1),y=c(0,1))

ggplot(Leic10_cal, aes(value.pred, value.obs))+
  geom_point(shape=15,size=3,colour="blue")+
  geom_abline(intercept = 0,slope=1,linetype="dashed")+
  scale_x_continuous(limits = c(0,0.3)) +
  scale_y_continuous(limits=c(0,0.3)) +
  labs(x="Predicted risk",y="Observed risk",title="")


```

```{r RP presentation Leicprob,eval=FALSE}

pred_obs<-Calibration_df(data=rs_train,predRisk="Leicrs_prob",groups=10,times=10,format="long")
  
#Plot observed vs predicted 
ggplot(pred_obs, aes(group, value, colour = variable)) + 
                  geom_point(size = 3, shape=19) + 
                  labs(title = "Calibration of the Leicester score at ten years using the training data.", x = "10th predicted risk", y = "10-year T2D risk") +
                  scale_color_manual(labels = c("Pred"="Predicted risk from Leicester score", "Obs"="Observed risk"), 
                                     values = c("Pred"="blue", "Obs"="red")) + 
                  theme(legend.title = element_blank())
```


# Enhanced score without PRS (Training)

We computed the predicted probability of developing T2D within 10 years specified below using the betas of the enhanced model (without PRS) and plotted them against observed 9-year risk.

```{r}

# Model 1: Ethnicity + other Leicester vars

RFs=c("TEU_BaC_AgeCat","BaC_Sex","TEU_ethnicgrp","TEU_FaH_diab","BSM_WaistCirc_Cat","BSM_BMI_Cat","HighBP")

formula<-as.formula(paste0("Surv(TEU_T2DM_time,TEU_T2DM_status)~",
                           paste(RFs, collapse="+")))


Eth_model<-coxph(formula,data = tot_train,x=TRUE)

# Create list of betas 
Eth_model_betas<-Eth_model$coefficients

Eth_model_betas<-data.frame(round(Eth_model_betas*10,0))

Eth_model_betas <- tibble::rownames_to_column(Eth_model_betas, "VALUE")
colnames(Eth_model_betas)<-c("VALUE","New Score 1")

# Present our model 
Eth_model_output<-printcoxresults(df=tot_train,varlist=RFs,modeloutput = Eth_model,pretty_names=pretty_names,forplot = FALSE,print_beta = TRUE,IDcol = TRUE)

```

```{r PRS model}

# Fit a model
vars=c(RFs,"PRS_quintiles","GeP_Array")

formula<-as.formula(paste0("Surv(TEU_T2DM_time,TEU_T2DM_status)~",
                           paste(vars, collapse="+")))


PRS_model<-coxph(formula,data = tot_train,x=TRUE)

```


```{r}
# Compute the predicted probability
rs_train<-rs_train%>%
  mutate(# Predicted 10-year risk score (RS)
    eth_prob10=1-predictCox(Eth_model,times=10,newdata = rs_train)$survival,
    eth_PRS_prob10=1-predictCox(PRS_model,times=10, newdata = rs_train)$survival)

```

Title: Calibration plot of the enhanced score (without PRS) with 10-year risk of T2D using the training data.

```{r}

enh_calib10<-Calibration_df(predRisk = "eth_prob10",data=rs_train,groups=5,times=9,format="wide")

ggplot(enh_calib10, aes(value.pred, value.obs))+
  geom_point(shape=15,size=3,colour="blue")+
  geom_abline(intercept = 0,slope=1,linetype="dashed")+
  scale_x_continuous(limits = c(0,0.3)) +
  scale_y_continuous(limits=c(0,0.3)) +
  labs(x="Predicted risk",y="Observed risk",title="")

```


# Enhanced score with PRS (Training)

We computed the predicted probability of developing T2D within 10 years using the betas of the enhanced model (with PRS) and plotted them against the observed 9-year risk.

Note: We incorporated the genetic array information when computing the predicted probability because I am not sure how to compute predicted probability regardless of the genetic array.

Title: Calibration plot of the enhanced score (with PRS) with 10-year risk of T2D using the training data

```{r}

enh_PRS_calib10<-Calibration_df(data=rs_train,predRisk = "eth_PRS_prob10",times=9,groups = 5,format = "wide")

ggplot(enh_PRS_calib10, aes(value.pred, value.obs))+
  geom_point(shape=15,size=3,colour="blue")+
  geom_abline(intercept = 0,slope=1,linetype="dashed")+
  scale_x_continuous(limits = c(0,0.3)) +
  scale_y_continuous(limits=c(0,0.3)) +
  labs(x="Predicted risk",y="Observed risk",title="")

```

# Leicester score (Test)

```{r pred prob test data}

tot_test<-tot_test%>%select(!starts_with("RS_"))

# Leicester risk score vars
columns=c("TEU_BaC_AgeCat","BaC_Sex","Ethnicity","TEU_FaH_diab","BSM_WaistCirc_Cat","BSM_BMI_Cat","HighBP")

rs_test<-cbind(tot_test, setNames(tot_test[columns], paste0("RS_", columns)))

rs_test<-rs_test%>%
  mutate_at(paste0("RS_",columns), ~var_to_beta(.))

# Sum all the columns starting with RS_

rs_test<-rs_test%>%
  mutate(sum_beta=-3.430668+rowSums(select(.,starts_with("RS_"))))%>%
  mutate(Leicrs_prob=exp(sum_beta)/(1+exp(sum_beta)))%>%
  mutate(# Predicted 10-year risk score (RS)
    eth_prob10=1-predictCox(Eth_model,times=10,newdata = tot_test)$survival,
    eth_PRS_prob10=1-predictCox(PRS_model,times=10, newdata = tot_test)$survival)

```

```{r}

rs_test<-rs_test%>%mutate(
    pred_groups=Percentile(Leicrs_prob,probs=seq(0, 1, 1/5),labels = c(1:5),right=TRUE))
  
survfit2(Surv(TEU_T2DM_time,TEU_T2DM_status) ~ pred_groups, data = rs_test) %>% 
  ggsurvfit() +
  labs(
    title = "KM plot by quintiles of predicted probability from the Leicester model \n using the test data",
    x = "Years",
    y = "Overall survival probability"
    ) + 
  add_confidence_interval() +
  scale_x_continuous(breaks = seq(0,12,2)) +
  add_risktable()

```

Title: Calibration plot of the Leicester score with 10-year risk of T2D using the test data.

```{r Leicprob calibration test}

Leic10_cal<-Calibration_df(data=rs_test,predRisk="Leicrs_prob",groups=5,times=9,format="wide")

#plot(pred_obs_wide$value.pred,pred_obs_wide$value.obs,main="",#xlab="",ylab="",
#pch=16,ps=2, xlim=c(0,0.4) , ylim=c(0,0.4),cex.lab=1.2,
#cex.axis=1.1, las=1)
#lines(x =c(0,1),y=c(0,1))

ggplot(Leic10_cal, aes(value.pred, value.obs))+
  geom_point(shape=15,size=3,colour="blue")+
  geom_abline(intercept = 0,slope=1,linetype="dashed")+
  scale_x_continuous(limits = c(0,0.3)) +
  scale_y_continuous(limits=c(0,0.3)) +
  labs(x="Predicted risk",y="Observed risk",title="")


```

# Enhanced score without PRS (Test)

Title: Calibration plot of the enhanced score (without PRS) with 10-year risk of T2D using the test data.

```{r}

enh_calib10<-Calibration_df(predRisk = "eth_prob10",data=rs_test,groups=5,times=9,format="wide")

ggplot(enh_calib10, aes(value.pred, value.obs))+
  geom_point(shape=15,size=3,colour="blue")+
  geom_abline(intercept = 0,slope=1,linetype="dashed")+
  scale_x_continuous(limits = c(0,0.3)) +
  scale_y_continuous(limits=c(0,0.3)) +
  labs(x="Predicted risk",y="Observed risk",title="")

```

# Enhanced score with PRS (Test)

```{r}

predRisk = "eth_PRS_prob10";times=10;groups = 5

rs_test<-rs_test%>%mutate(
     pred_deciles=Percentile((!!sym(predRisk)),probs=seq(0, 1, 1/groups),labels = c(1:groups),right=TRUE))

survfit2(Surv(TEU_T2DM_time,TEU_T2DM_status) ~ pred_deciles, data = rs_test) %>% 
  ggsurvfit() +
  labs(
    x = "Years",
    y = "Overall survival probability"
    ) + 
  add_confidence_interval() +
  scale_x_continuous(breaks = seq(0,12,2)) +
  add_risktable()

```
Title: Calibration plot of the enhanced score (with PRS) with 10-year risk of T2D using the test data

```{r}

enh_PRS_calib10<-Calibration_df(data=rs_test,predRisk = "eth_PRS_prob10",times=9,groups = 5,format = "wide")

ggplot(enh_PRS_calib10, aes(value.pred, value.obs))+
  geom_point(shape=15,size=3,colour="blue")+
  geom_abline(intercept = 0,slope=1,linetype="dashed")+
  scale_x_continuous(limits = c(0,0.3)) +
  scale_y_continuous(limits=c(0,0.3)) +
  labs(x="Predicted risk",y="Observed risk",title="")

```


# OLD 

We assessed the calibration of Leicester score, two enhanced scores using the training data. 

The original Leicester score is not in the probability scale, but the [diabetes website](https://riskscore.diabetes.org.uk/start) provides the predicted probability of developing T2D within 10 years for each risk group (copied below):

*	Low 0-6 risk category: 1 in 100 people with your risk will get T2D in the next 10 years. (p=0.01)
*	Increased 7-15: 1 in 35. (p=0.03)
* Moderate: 16-24: 1 in 10. (p=0.1)
* High 25-47: 1 in 4. (p=0.25)

We used the probabilities above as the predicted probability (plotted in blue dots) and computed the cumulative incidence via 1- Kaplan Meier estimate as the observed probability (plotted in red dots).


```{r webpage,eval=FALSE}

rs_train<-rs_train%>%mutate(group=as.numeric(diab_rs_cat))
  
# Predicted risk
pred<-tibble(group=1:4,
             value=case_when(group==1 ~ 1/100,
                           group==2 ~ 1/35,
                           group==3 ~ 1/10,
                           group==4 ~ 1/4),
           variable="Pred")


# Observed onset risk in each risk group
fit<-survfit(Surv(TEU_T2DM_time,TEU_T2DM_status)~group,data=rs_train)
#res.sum <- surv_summary(fit)

sum10<-summary(fit,times = 10)

sum10_df <- data.frame(group=sum10$strata,
                  surv = sum10$surv
                  )

obs<-sum10_df%>%
  mutate(group=as.numeric(str_remove(group,"group=")),
         value=1-surv,
           variable="Obs")%>%
  select(group,variable,value)
  
# Combine
pred_obs<-rbind(pred,obs)# %>%
    #group_by(group, variable) %>%
    #summarise(value = mean(value), .groups='drop')%>%
  #mutate(group=as.numeric(group),variable=as.factor(variable))
  
#Plot observed vs predicted 
ggplot(pred_obs, aes(group, value, colour = variable)) + 
                  geom_point(size = 3, shape=19) + 
                  labs(title = "Calibration of the Leicester score at ten years using the training data.", x = "Risk group", y = "10-year T2D risk") +
                  scale_color_manual(labels = c("Pred"="Predicted risk from Leicester score", "Obs"="Observed risk"), 
                                     values = c("Pred"="blue", "Obs"="red")) + 
                  theme(legend.title = element_blank()) + 
                  scale_x_continuous(breaks = c(1, 2, 3, 4),
                                     labels = c("Low:0-6", "Increased:7-15", "Moderate:16-24", "High:25-47")) 

```


We then produced the calibration plots for the two enhanced scores. The differences here were that we computed the predicted 10-year T2D probability for each participant using the Cox models and took the mean within each risk group (plotted in blue). The observed risk followed the same approach as above.

```{r,eval=FALSE}

# Model 1: Ethnicity + other Leicester vars

RFs=c("TEU_BaC_AgeCat","BaC_Sex","TEU_ethnicgrp","TEU_FaH_diab","BSM_WaistCirc_Cat","BSM_BMI_Cat","HighBP")

formula<-as.formula(paste0("Surv(TEU_T2DM_time,TEU_T2DM_status)~",
                           paste(RFs, collapse="+")))


Eth_model<-coxph(formula,data = tot_train,x=TRUE)

# Create list of betas 
Eth_model_betas<-Eth_model$coefficients

Eth_model_betas<-data.frame(round(Eth_model_betas*10,0))

Eth_model_betas <- tibble::rownames_to_column(Eth_model_betas, "VALUE")
colnames(Eth_model_betas)<-c("VALUE","New Score 1")

# Present our model 
Eth_model_output<-printcoxresults(df=tot_train,varlist=RFs,modeloutput = Eth_model,pretty_names=pretty_names,forplot = FALSE,print_beta = TRUE,IDcol = TRUE)

```

```{r,eval=FALSE}

# Fit a model
vars=c(RFs,"PRS_quintiles","GeP_Array")

formula<-as.formula(paste0("Surv(TEU_T2DM_time,TEU_T2DM_status)~",
                           paste(vars, collapse="+")))


PRS_model<-coxph(formula,data = tot_train,x=TRUE)

```

```{r present Eth and PRS model,eval=FALSE}

#Create list of new betas
PRS_model_betas<-PRS_model$coefficients

# multiply all betas by 10 and round to the nearest integer
PRS_model_betas<-data.frame(round(PRS_model_betas*10,0))

PRS_model_betas <- tibble::rownames_to_column(PRS_model_betas, "VALUE")
colnames(PRS_model_betas)<-c("VALUE","New Score 2")


# Present the model 
PRS_model_output<-printcoxresults(df=tot_train,varlist=c(RFs,"PRS_quintiles"),modeloutput = PRS_model,pretty_names=pretty_names,forplot = FALSE,print_beta = TRUE,IDcol = TRUE)

# Processing 
model_output<-Eth_model_output%>%
  # add new score 1 
  left_join(.,Eth_model_betas,by=c("IDcol"="VALUE"))%>%
  mutate(`New Score 1`=replace_na(`New Score 1`,0))%>%
  # Full join the PRS model
  full_join(.,PRS_model_output,by="IDcol")%>%
  # Add new score 2
  left_join(.,PRS_model_betas,by=c("IDcol"="VALUE"))%>%
  mutate(`New Score 2`=replace_na(`New Score 2`,0))%>%
  select(Coefficient.y,Levels.y,Beta.x,`95% CI.x`,p.x,`New Score 1`,Beta.y,`95% CI.y`,`p.y`,`New Score 2`)


# Eth model
Eth_betas_list<-setNames(na.omit(model_output$`New Score 1`),model_output%>%filter(!is.na(`New Score 1`))%>%pull(Levels.y))

# new betas list
PRS_betas_list <- setNames(model_output$`New Score 2`,model_output$Levels.y)

```

```{r compute new risk scores training,eval=FALSE}

# Prep for computing risk scores
columns=c("TEU_BaC_AgeCat","BaC_Sex","TEU_ethnicgrp","TEU_FaH_diab","BSM_WaistCirc_Cat","BSM_BMI_Cat","HighBP")

columns_PRS=c(columns,"PRS_quintiles")

# Make duplicate columns for variables to be transferred to corresponding scores

rs_train<-cbind(tot_train,
                setNames(tot_train[columns], paste0("Eth_", columns)),
                setNames(tot_train[columns_PRS], paste0("NRSPRS_", columns_PRS)))

# 2. Compute the risk score
#https://dplyr.tidyverse.org/reference/recode.html#ref-usage
rs_train<-rs_train%>%
  # Compute enhanced risk score 1
  mutate_at(paste0("Eth_",columns), ~recode(.,!!!Eth_betas_list))%>%
  mutate(Eth_rs=rowSums(select(.,starts_with("Eth_"))))%>%
  # Compute enhanced risk score 2
  # Essential to add .default=0 because the beta list doesn't have all possible levels (e.g. 40-49 Yes high BP)
  mutate_at(paste0("NRSPRS_",columns_PRS), ~recode(.,!!!PRS_betas_list,.default = 0))%>%
  mutate(PRS_rs=rowSums(select(.,starts_with("NRSPRS_"))))%>%
  # create status at 10 years
  mutate(T2D10_status=ifelse(TEU_T2DM_status==1 & TEU_T2DM_time<=10,1,0))

```

```{r compute new risk scores test,eval=FALSE}

# Make duplicate columns for variables to be transferred to corresponding scores

rs_test<-cbind(tot_test,
               setNames(tot_test[columns], paste0("Eth_", columns)),
               setNames(tot_test[columns_PRS], paste0("NRSPRS_", columns_PRS)))

# 2. Compute the risk score using age int scoring
#https://dplyr.tidyverse.org/reference/recode.html#ref-usage
rs_test<-rs_test%>%
  # Compute enhanced risk score 1
  mutate_at(paste0("Eth_",columns), ~recode(.,!!!Eth_betas_list))%>%
  mutate(Eth_rs=rowSums(select(.,starts_with("Eth_"))))%>%
  # Compute enhanced risk score 2
  # Essential to add .default=0 because the beta list doesn't have all possible levels (e.g. 40-49 Yes high BP)
  mutate_at(paste0("NRSPRS_",columns_PRS), ~recode(.,!!!PRS_betas_list,.default = 0))%>%
  mutate(PRS_rs=rowSums(select(.,starts_with("NRSPRS_"))))%>%
  # create status at 10 years
  mutate(T2D10_status=ifelse(TEU_T2DM_status==1 & TEU_T2DM_time<=10,1,0))

```

```{r optimal cutoffs,eval=FALSE}

# 60% seems a good fit
Leicrs16_perc=pretty_dp(rs_train%>%filter(diab_rs<16)%>%nrow()/nrow(rs_train),pct=TRUE,dp=0)

Eth_threshold=quantile(rs_train$Eth_rs,as.numeric(Leicrs16_perc)/100)

PRS_threshold=quantile(rs_train$PRS_rs,as.numeric(Leicrs16_perc)/100) 

```


```{r compute 10-year risk of enhanced models,eval=FALSE}
# Categorise individuals into low and high risk based on enhanced scores
rs_train$Eth_rs_binary<-ifelse(rs_train$Eth_rs<Eth_threshold,1,2)
rs_train$PRS_rs_binary<-ifelse(rs_train$PRS_rs<PRS_threshold,1,2)

# Compute the predicted probability
rs_train<-rs_train%>%
  mutate(# Predicted 10-year risk score (RS)
    eth_rs=1-predictCox(Eth_model,times=10,newdata = rs_train)$survival,
    eth_PRS_rs=1-predictCox(PRS_model,times=10, newdata = rs_train)$survival)

pred<-rs_train%>%
  group_by(Eth_rs_binary)%>%
  summarise(value=mean(eth_rs))%>%
  mutate(variable="Pred")%>%rename(group=Eth_rs_binary)

# Observed onset risk in each risk group
fit<-survfit(Surv(TEU_T2DM_time,TEU_T2DM_status)~Eth_rs_binary,data=rs_train)
#res.sum <- surv_summary(fit)

sum10<-summary(fit,times = 10)

sum10_df <- data.frame(group=sum10$strata,
                  surv = sum10$surv
                  )

obs<-sum10_df%>%
  mutate(group=as.numeric(str_remove(group,"Eth_rs_binary=")),
         value=1-surv,
           variable="Obs")%>%
  select(group,variable,value)
  
# Combine
pred_obs<-rbind(pred,obs)# %>%
    #group_by(group, variable) %>%
    #summarise(value = mean(value), .groups='drop')%>%
  #mutate(group=as.numeric(group),variable=as.factor(variable))
  
#Plot observed vs predicted 
ggplot(pred_obs, aes(group, value, colour = variable)) + 
                  geom_point(size = 3, shape=19) + 
                  labs(title = "Calibration of the enhanced score (without PRS) at ten years using the training data.", x = "Risk group", y = "10-year T2D risk") +
                  scale_color_manual(labels = c("Pred"="Predicted risk from enhanced score (without PRS)", "Obs"="Observed risk"), 
                                     values = c("Pred"="blue", "Obs"="red")) + 
                  theme(legend.title = element_blank()) + 
                  scale_x_continuous(breaks = c(1, 2),
                                     labels = c("Low:0-18", "High:19-59")) 
```

```{r,eval=FALSE}

pred<-rs_train%>%
  group_by(PRS_rs_binary)%>%
  summarise(value=mean(eth_PRS_rs))%>%
  mutate(variable="Pred")%>%rename(group=PRS_rs_binary)

# Observed onset risk in each risk group
fit<-survfit(Surv(TEU_T2DM_time,TEU_T2DM_status)~PRS_rs_binary,data=rs_train)
#res.sum <- surv_summary(fit)

sum10<-summary(fit,times = 10)

sum10_df <- data.frame(group=sum10$strata,
                  surv = sum10$surv
                  )

obs<-sum10_df%>%
  mutate(group=as.numeric(str_remove(group,"PRS_rs_binary=")),
         value=1-surv,
           variable="Obs")%>%
  select(group,variable,value)
  
# Combine
pred_obs<-rbind(pred,obs)# %>%
    #group_by(group, variable) %>%
    #summarise(value = mean(value), .groups='drop')%>%
  #mutate(group=as.numeric(group),variable=as.factor(variable))
  
#Plot observed vs predicted 
ggplot(pred_obs, aes(group, value, colour = variable)) + 
                  geom_point(size = 3, shape=19) + 
                  labs(title = "Calibration of the enhanced score (with PRS) at ten years using the training data.", x = "Risk group", y = "10-year T2D risk") +
                  scale_color_manual(labels = c("Pred"="Predicted risk from enhanced score (with PRS)", "Obs"="Observed risk"), 
                                     values = c("Pred"="blue", "Obs"="red")) + 
                  theme(legend.title = element_blank()) + 
                  scale_x_continuous(breaks = c(1, 2),
                                     labels = c("Low:0-27", "High:28-73")) 
```


# Calibration using test data

We followed the same steps as above but with test data instead.

```{r,eval=FALSE}

rs_test<-rs_test%>%mutate(group=as.numeric(diab_rs_cat))
  
# Predicted risk
pred<-tibble(group=1:4,
             value=case_when(group==1 ~ 1/100,
                           group==2 ~ 1/35,
                           group==3 ~ 1/10,
                           group==4 ~ 1/4),
           variable="Pred")


# Observed onset risk in each risk group
fit<-survfit(Surv(TEU_T2DM_time,TEU_T2DM_status)~group,data=rs_test)
#res.sum <- surv_summary(fit)

sum10<-summary(fit,times = 10)

sum10_df <- data.frame(group=sum10$strata,
                  surv = sum10$surv
                  )

obs<-sum10_df%>%
  mutate(group=as.numeric(str_remove(group,"group=")),
         value=1-surv,
           variable="Obs")%>%
  select(group,variable,value)
  
# Combine
pred_obs<-rbind(pred,obs)# %>%
    #group_by(group, variable) %>%
    #summarise(value = mean(value), .groups='drop')%>%
  #mutate(group=as.numeric(group),variable=as.factor(variable))
  
#Plot observed vs predicted 
ggplot(pred_obs, aes(group, value, colour = variable)) + 
                  geom_point(size = 3, shape=19) + 
                  labs(title = "Calibration of the Leicester score at ten years using the test data.", x = "Risk group", y = "10-year T2D risk") +
                  scale_color_manual(labels = c("Pred"="Predicted risk from Leicester score", "Obs"="Observed risk"), 
                                     values = c("Pred"="blue", "Obs"="red")) + 
                  theme(legend.title = element_blank()) + 
                  scale_x_continuous(breaks = c(1, 2, 3, 4),
                                     labels = c("Low:0-6", "Increased:7-15", "Moderate:16-24", "High:25-47")) 

```

```{r,eval=FALSE}
# Categorise individuals into low and high risk based on enhanced scores
rs_test$Eth_rs_binary<-ifelse(rs_test$Eth_rs<Eth_threshold,1,2)
rs_test$PRS_rs_binary<-ifelse(rs_test$PRS_rs<PRS_threshold,1,2)

# Compute the predicted probability
rs_test<-rs_test%>%
  mutate(# Predicted 10-year risk score (RS)
    eth_rs=1-predictCox(Eth_model,times=10,newdata = rs_test)$survival,
    eth_PRS_rs=1-predictCox(PRS_model,times=10, newdata = rs_test)$survival)

pred<-rs_test%>%
  group_by(Eth_rs_binary)%>%
  summarise(value=mean(eth_rs))%>%
  mutate(variable="Pred")%>%rename(group=Eth_rs_binary)

# Observed onset risk in each risk group
fit<-survfit(Surv(TEU_T2DM_time,TEU_T2DM_status)~Eth_rs_binary,data=rs_test)
#res.sum <- surv_summary(fit)

sum10<-summary(fit,times = 10)

sum10_df <- data.frame(group=sum10$strata,
                  surv = sum10$surv
                  )

obs<-sum10_df%>%
  mutate(group=as.numeric(str_remove(group,"Eth_rs_binary=")),
         value=1-surv,
           variable="Obs")%>%
  select(group,variable,value)
  
# Combine
pred_obs<-rbind(pred,obs)# %>%
    #group_by(group, variable) %>%
    #summarise(value = mean(value), .groups='drop')%>%
  #mutate(group=as.numeric(group),variable=as.factor(variable))
  
#Plot observed vs predicted 
ggplot(pred_obs, aes(group, value, colour = variable)) + 
                  geom_point(size = 3, shape=19) + 
                  labs(title = "Calibration of the enhanced score (without PRS) at ten years using the test data.", x = "Risk group", y = "10-year T2D risk") +
                  scale_color_manual(labels = c("Pred"="Predicted risk from enhanced score (without PRS)", "Obs"="Observed risk"), 
                                     values = c("Pred"="blue", "Obs"="red")) + 
                  theme(legend.title = element_blank()) + 
                  scale_x_continuous(breaks = c(1, 2),
                                     labels = c("Low:0-18", "High:19-59"))
```

```{r,eval=FALSE}

pred<-rs_test%>%
  group_by(PRS_rs_binary)%>%
  summarise(value=mean(eth_PRS_rs))%>%
  mutate(variable="Pred")%>%rename(group=PRS_rs_binary)

# Observed onset risk in each risk group
fit<-survfit(Surv(TEU_T2DM_time,TEU_T2DM_status)~PRS_rs_binary,data=rs_test)
#res.sum <- surv_summary(fit)

sum10<-summary(fit,times = 10)

sum10_df <- data.frame(group=sum10$strata,
                  surv = sum10$surv
                  )

obs<-sum10_df%>%
  mutate(group=as.numeric(str_remove(group,"PRS_rs_binary=")),
         value=1-surv,
           variable="Obs")%>%
  select(group,variable,value)
  
# Combine
pred_obs<-rbind(pred,obs)# %>%
    #group_by(group, variable) %>%
    #summarise(value = mean(value), .groups='drop')%>%
  #mutate(group=as.numeric(group),variable=as.factor(variable))
  
#Plot observed vs predicted 
ggplot(pred_obs, aes(group, value, colour = variable)) + 
                  geom_point(size = 3, shape=19) + 
                  labs(title = "Calibration of the enhanced score (with PRS) at ten years using the test data.", x = "Risk group", y = "10-year T2D risk") +
                  scale_color_manual(labels = c("Pred"="Predicted risk from enhanced score (with PRS)", "Obs"="Observed risk"), 
                                     values = c("Pred"="blue", "Obs"="red")) + 
                  theme(legend.title = element_blank()) + 
                  scale_x_continuous(breaks = c(1, 2),
                                     labels = c("Low:0-27", "High:28-73")) 
```











