---
title: 'SA: Further exclude gestational diabetes'
author: "Xiaonan Liu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  word_document:
      fig_width: 8
      fig_height: 5
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

# Introduction

In this senstivity analysis, we will further exclude prevalent gestational diabetes and rerun the main analyses (i.e. model fitting, AUC, NRI).

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(kableExtra)
library(knitr)
library(yaml)
library(here)
library(ggplot2)
library(tidyr)
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(svglite)
library(rsvg)
library(png)
library(dplyr)
library(survival)
library(survminer)
library(Rcpp)
library(summarytools)
library(arsenal)
library(pander)
library(ggeffects)
library(sjPlot)
library(etm)
library(readr)
library(tibble)
library(ResourceSelection)
library(Publish)
library(rsample)
library(plotROC)
library(pROC)
library(nricens)
library(PredictABEL)
library(cutpointr)
library(caret)
library(gridExtra)
# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
#options(knitr.kable.NA='', knitr.table.format = "html")
#options(kableExtra.auto_format = FALSE)
# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('missing', '')


knitr::opts_knit$set(root.dir = here::here())

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(config$functions))
source(here::here(config$scripts$nri))

my_controls <- tableby.control(
    test = F,
    total = T,
    #numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss"), 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      medianq1q3 = "Median (IQR)",
      range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )

# Read in data derived from code chunk below
data<-readRDS(here::here(file.path(config$data$derived,"data_Leicrs.RDS")))

tot_train<-readRDS(here::here(file.path(config$data$derived,"tot_train.rds")))
tot_test<-readRDS(here::here(file.path(config$data$derived,"tot_test.rds")))


```

```{r,include=FALSE}
source(file.path(config$scripts$cleaning, "dataset_generator.R"))

pretty_func <- pretty_switch(field_definitions=TEU_SPECS$T2DM_PRS, return_type = "function")
pretty_names <- pretty_switch(field_definitions=TEU_SPECS$T2DM_PRS, return_type="list")

pretty_names$BaC_Sex="Sex"
pretty_names$hba1c_cat="HbA1c at baseline"
pretty_names$hba1c_cat2="HbA1c at baseline"

pretty_names$BSM_WaistCirc_Cat="Waist circumference,cm"
pretty_names$BSM_BMI_Cat="BMI"
pretty_names$HighBP="Antihypertensive meds or high blood pressure"
pretty_names$diab_rs_cat="Leicester risk score"
pretty_names$PRS_quintiles="T2D PRS quintiles"
pretty_names$Ethnicity="Ethnicity"

pretty_names$diab_rs="LRA Score"
pretty_names$Eth_rs="LRArev Score"
pretty_names$PRS_rs="LRAprs Score"
```

In main analysis, the sample size of training data is N= `r nrow(tot_train)` and that of test data is N=`r nrow(tot_test)`.

```{r excl gestational}

tot_train<-tot_train%>%filter(Estw_diab_prevalent=="Diabetes unlikely")
tot_test<-tot_test%>%filter(Estw_diab_prevalent=="Diabetes unlikely")

```

After excluding prevalent gestational diabetes at baseline, the sample size of training data is N= `r nrow(tot_train)` and that of test data is N=`r nrow(tot_test)`.

# Analyses

In summary, we followed the steps below: 

* Build two separate models:

  - The enhanced model without PRS
  - The enhanced model with PRS

* Calculate three sets of risk scores for individuals: the original Leicester score, the enhanced score without PRS, the enhanced score with PRS.

* Compare model performance among all three risk scores with onset T2D.


## Model building 

Note: We only computed the Leicester score for the study population using their published scoring system. We did NOT build a separate model fitting the original Leicester variables. 

We built two separate models: 

**Model 1: The enhanced model without PRS**:

A Cox model: Incident T2D ~ Age + Sex + Ethnicity (more refined groupings)* + FaH of diab + Waist circumference + BMI + High BP

Note*: The more refined groupings of ethnicity are White (ref), Black, South Asian, Mixed, Other. The ethnicity variable in the original Leicester score is binary with the groupings: White (ref) and Other.

```{r Build the enhanced model 1}
# Model 1: Ethnicity + other Leicester vars

RFs=c("TEU_BaC_AgeCat","BaC_Sex","TEU_ethnicgrp","TEU_FaH_diab","BSM_WaistCirc_Cat","BSM_BMI_Cat","HighBP")

formula<-as.formula(paste0("Surv(TEU_T2DM_time,TEU_T2DM_status)~",
                           paste(RFs, collapse="+")))


Eth_model<-coxph(formula,data = tot_train)

# Create list of betas 
Eth_model_betas<-Eth_model$coefficients

Eth_model_betas<-data.frame(round(Eth_model_betas*10,0))

Eth_model_betas <- tibble::rownames_to_column(Eth_model_betas, "VALUE")
colnames(Eth_model_betas)<-c("VALUE","New Score 1")

# Present our model 
Eth_model_output<-printcoxresults(df=tot_train,varlist=RFs,modeloutput = Eth_model,pretty_names=pretty_names,forplot = FALSE,print_beta = TRUE,IDcol = TRUE)

# Processing 

#Eth_model_output<-Eth_model_output%>%
  # add Leicester risk score 
  #mutate(`Leicester score`=var_to_score(Levels))%>%
  # add new score 
  #left_join(.,Eth_model_betas,by=c("IDcol"="VALUE"))%>%
  #mutate(`New Score 1`=replace_na(`New Score 1`,0))

#pander(Eth_model_output,caption = "The model output of the enhanced Leicester model, replacing the original binary ethnicity variable with a more refined ethnicity variable.")

```

**Model 2: The enhanced PRS model**: Further adding PRS (in quintiles) to the model 1 above:

A Cox model: Incident T2D status ~ Age + Sex + Ethnicity (more refined groupings) + FaH of diab + Waist circumference + BMI + High BP + PRS (quintiles) + genetic array.

Note: genetic array was included in the model only for adjustment. Its beta was not used to construct a risk score.


```{r PRS model}

# Fit a model
vars=c(RFs,"PRS_quintiles","GeP_Array")

formula<-as.formula(paste0("Surv(TEU_T2DM_time,TEU_T2DM_status)~",
                           paste(vars, collapse="+")))


PRS_model<-coxph(formula,data = tot_train)

```


```{r present Eth and PRS model}

#Create list of new betas
PRS_model_betas<-PRS_model$coefficients

# multiply all betas by 10 and round to the nearest integer
PRS_model_betas<-data.frame(round(PRS_model_betas*10,0))

PRS_model_betas <- tibble::rownames_to_column(PRS_model_betas, "VALUE")
colnames(PRS_model_betas)<-c("VALUE","New Score 2")


# Present the model 
PRS_model_output<-printcoxresults(df=tot_train,varlist=c(RFs,"PRS_quintiles"),modeloutput = PRS_model,pretty_names=pretty_names,forplot = FALSE,print_beta = TRUE,IDcol = TRUE)

# Processing 
model_output<-Eth_model_output%>%
  # add new score 1 
  left_join(.,Eth_model_betas,by=c("IDcol"="VALUE"))%>%
  mutate(`New Score 1`=replace_na(`New Score 1`,0))%>%
  # Full join the PRS model
  full_join(.,PRS_model_output,by="IDcol")%>%
  # Add new score 2
  left_join(.,PRS_model_betas,by=c("IDcol"="VALUE"))%>%
  mutate(`New Score 2`=replace_na(`New Score 2`,0))%>%
  select(Coefficient.y,Levels.y,Beta.x,`95% CI (Beta).x`,HR.x,`95% CI (HR).x`,p.x,`New Score 1`,Beta.y,`95% CI (Beta).y`,HR.y,`95% CI (HR).y`,`p.y`,`New Score 2`)


# Eth model
Eth_betas_list<-setNames(na.omit(model_output$`New Score 1`),model_output%>%filter(!is.na(`New Score 1`))%>%pull(Levels.y))

# new betas list
PRS_betas_list <- setNames(model_output$`New Score 2`,model_output$Levels.y)

```

```{r pretty present models}

# for presenting only
Eth_temp<-printcoxresults(df=tot_train,varlist=RFs,modeloutput = Eth_model,pretty_names=pretty_names,forplot = FALSE,print_beta = TRUE,IDcol = TRUE,onecol = TRUE)

PRS_temp<-printcoxresults(df=tot_train,varlist=c(RFs,"PRS_quintiles"),modeloutput = PRS_model,pretty_names=pretty_names,forplot = FALSE,print_beta = TRUE,IDcol = TRUE,onecol = TRUE)

# Processing 
model_temp<-Eth_temp%>%
  # add new score 1 
  left_join(.,Eth_model_betas,by=c("IDcol"="VALUE"))%>%
  # Full join the PRS model
  full_join(.,PRS_temp,by="IDcol")%>%
  # Add new score 2
  left_join(.,PRS_model_betas,by=c("IDcol"="VALUE"))%>%
  mutate(
    `New Score 1`=ifelse(Beta.x=="Reference",0,`New Score 1`),
    `New Score 2`=ifelse(Beta.y=="Reference",0,`New Score 2`))%>%
  mutate_if(is.character, ~replace_na(.,""))%>%
  mutate(Beta_CI.x=paste0(Beta.x," ",`95% CI (Beta).x`),
         Beta_CI.y=paste0(Beta.y," ",`95% CI (Beta).y`),
         HR_CI.x=paste0(HR.x," ",`95% CI (HR).x`),
         HR_CI.y=paste0(HR.y," ",`95% CI (HR).y`))%>%
  select(Coefficient.y,`Beta_CI.x`,`HR_CI.x`,p.x,`New Score 1`,`Beta_CI.y`,`HR_CI.y`,`p.y`,`New Score 2`)


pander(model_temp,caption = "Model outputs and scoring system of the enhanced score without PRS model and with PRS model, derived using the training data. 'New score 1' was derived from the enhanced model without PRS and 'New score 2' was derived from the enhanced model with PRS. Both new scores were derived by multiplying the beta coefficients from the corresponding model by 10 and rounding up to the nearest integer.")

```

To compute the new risk score from the two enhanced models, we followed the same approach as the Leicester risk score: Multiply the beta coefficients from each model by 10 and round up to the nearest integer (See 'New Score 1' and 'New Score 2' above). These scores are then summed to give the total score.

```{r compute new risk scores training}

# Prep for computing risk scores
columns=c("TEU_BaC_AgeCat","BaC_Sex","TEU_ethnicgrp","TEU_FaH_diab","BSM_WaistCirc_Cat","BSM_BMI_Cat","HighBP")

columns_PRS=c(columns,"PRS_quintiles")

# Make duplicate columns for variables to be transferred to corresponding scores

rs_train<-cbind(tot_train,
                setNames(tot_train[columns], paste0("Eth_", columns)),
                setNames(tot_train[columns_PRS], paste0("NRSPRS_", columns_PRS)))

# 2. Compute the risk score
#https://dplyr.tidyverse.org/reference/recode.html#ref-usage
rs_train<-rs_train%>%
  # Compute enhanced risk score 1
  mutate_at(paste0("Eth_",columns), ~recode(.,!!!Eth_betas_list))%>%
  mutate(Eth_rs=rowSums(select(.,starts_with("Eth_"))))%>%
  # Compute enhanced risk score 2
  # Essential to add .default=0 because the beta list doesn't have all possible levels (e.g. 40-49 Yes high BP)
  mutate_at(paste0("NRSPRS_",columns_PRS), ~recode(.,!!!PRS_betas_list,.default = 0))%>%
  mutate(PRS_rs=rowSums(select(.,starts_with("NRSPRS_"))))%>%
  # create status at 10 years
  mutate(T2D10_status=ifelse(TEU_T2DM_status==1 & TEU_T2DM_time<=10,1,0))

```


```{r compute new risk scores test}

# Make duplicate columns for variables to be transferred to corresponding scores

rs_test<-cbind(tot_test,
               setNames(tot_test[columns], paste0("Eth_", columns)),
               setNames(tot_test[columns_PRS], paste0("NRSPRS_", columns_PRS)))

# 2. Compute the risk score using age int scoring
#https://dplyr.tidyverse.org/reference/recode.html#ref-usage
rs_test<-rs_test%>%
  # Compute enhanced risk score 1
  mutate_at(paste0("Eth_",columns), ~recode(.,!!!Eth_betas_list))%>%
  mutate(Eth_rs=rowSums(select(.,starts_with("Eth_"))))%>%
  # Compute enhanced risk score 2
  # Essential to add .default=0 because the beta list doesn't have all possible levels (e.g. 40-49 Yes high BP)
  mutate_at(paste0("NRSPRS_",columns_PRS), ~recode(.,!!!PRS_betas_list,.default = 0))%>%
  mutate(PRS_rs=rowSums(select(.,starts_with("NRSPRS_"))))%>%
  # create status at 10 years
  mutate(T2D10_status=ifelse(TEU_T2DM_status==1 & TEU_T2DM_time<=10,1,0))

```

```{r optimal cutoffs}
# Use cutpointr package to determine the optimal cut-off point for our new risk score

# Eth model
#Eth_threshold<-cutpointr(rs_train,Eth_rs,TEU_T2DM_status,method=maximize_metric,metric=sum_sens_spec)

#PRS_threshold<-cutpointr(rs_train,PRS_rs,TEU_T2DM_status,method=maximize_metric,metric=sum_sens_spec)

# Choose threshold based on balance of Sens and Spec, prioritise Sens
# 60% seems a good fit
Leicrs16_perc=round((rs_train%>%filter(diab_rs<16)%>%nrow()/nrow(rs_train))*100,0)

Eth_threshold=quantile(rs_train$Eth_rs,Leicrs16_perc/100)

PRS_threshold=quantile(rs_train$PRS_rs,Leicrs16_perc/100) 

```

**Choosing a cutoff**: We used 16 as the cutoff values for the Leicester score. Cutoff 16 corresponds to `r Leicrs16_perc` percentile of the Leicester risk score (i.e. what percentage of people had their Leicester score below 16.).

We used this percentile as the cutoffs for the other two enhanced scores. Hence, the cutoff for the enhanced Leicester score (without PRS) is `r Eth_threshold` and the cutoff for the enhanced Leicester score (with PRS) is `r PRS_threshold`. 
We found those scores at `r Leicrs16_perc` percentile are reasonable cutoffs based on the sensitivity and specificity.

[Previous approach]: Previously we chose the cutoff based on maximising the sum of sensitivity and specificity. However, we found this approach led to lower sensitivity than that of the Leicester score.

```{r summary stats new score}

long_train<-rs_train%>%
  select(ID,TEU_T2DM_status,T2D10_status,diab_rs,Eth_rs,PRS_rs)%>%
  pivot_longer(!c("ID","TEU_T2DM_status","T2D10_status"), names_to = "Score", values_to = "Value")%>%
  mutate(Score=factor(Score,levels = c("diab_rs","Eth_rs","PRS_rs")))


```

```{r Create annotation for histograms,message=FALSE}

# Categorise individuals to low (1) and high risk (2) groups
rs_train$diab_rs_binary<-ifelse(rs_train$diab_rs<16,1,2)
rs_train$Eth_rs_binary<-ifelse(rs_train$Eth_rs<Eth_threshold,1,2)
rs_train$PRS_rs_binary<-ifelse(rs_train$PRS_rs<PRS_threshold,1,2)

```

```{r summary stats new score test}

# Histogtam 
long_test<-rs_test%>%
  select(ID,TEU_T2DM_status,T2D10_status,diab_rs,Eth_rs,PRS_rs)%>%
  pivot_longer(!c("ID","TEU_T2DM_status","T2D10_status"), names_to = "Score", values_to = "Value")%>%
  mutate(Score=factor(Score,levels = c("diab_rs","Eth_rs","PRS_rs")))


```

```{r Create annotation for histograms test,message=FALSE}

# Categorise individuals to low (1) and high risk (2) groups
rs_test$diab_rs_binary<-ifelse(rs_test$diab_rs<16,1,2)
rs_test$Eth_rs_binary<-ifelse(rs_test$Eth_rs<Eth_threshold,1,2)
rs_test$PRS_rs_binary<-ifelse(rs_test$PRS_rs<PRS_threshold,1,2)

```

## Model performance in the training data

### AUC

We assess the outcome of T2D diagnosed within 10 years. Title: ROC Curve for risk scores using training data N=`r nrow(tot_train)`.

```{r AUC training CI}

AUC_training<-sapply(c("diab_rs","Eth_rs","PRS_rs"),function(i) AUC_CI(df=rs_train,score=i,status = "T2D10_status"))


AUC_training_label=c(
  paste0("LRA Score: ",AUC_training[1]),
  paste0("LRArev Score: ",AUC_training[2]),
  paste0("LRAprs Score: ",AUC_training[3])
)


# Try different format presenting ROC curve from below
ggplot(long_train,aes(d=T2D10_status, m=Value,color=Score))+geom_roc(n.cuts = 0,linealpha = 1,size=1)+
  style_roc()+
  #scale_color_manual(values=cbPalette)
  scale_color_manual(name="AUC (95% CI)",labels =  AUC_training_label,values = c("#F8766D","#00BA38", "#6193FF")) +
  labs(title ="",caption = "")

```


### NRI

We have three scores: 1) the Leicester scores, 2) the enhanced scores (without PRS), and 3) the enhanced scores (with PRS). Since the NRI is for pair-wise comparison between scores, we computed the NRI for the following 3 comparisons:

* Leicester scores vs enhanced scores(without PRS)
* Enhanced scores (without PRS) vs enhanced scores (with PRS)
* Leicester scores vs enhanced scores (with PRS)


**Comparison 1: Leicester scores vs enhanced scores (without PRS)**:

* Step 1: Categorise individuals based on the Leicester risk scores to two risk groups: "Low: <16" and "High: $\ge$ 16". Note: 16 is the established risk threshold for the Leicester risk score.

* Step 2: Categorise individuals based on the enhanced scores (without PRS) to two risk groups: "Low:< `r Eth_threshold`" and "High: $\ge$ `r Eth_threshold`".

* Step 3: Compute the categorical NRI with the incident T2D within 10 years. 


```{r NRI training,echo=FALSE}

# reclassification table
reclassification_table(rs_train, match("T2D10_status",names(rs_train)), rs_train$diab_rs_binary, rs_train$Eth_rs_binary, cutoff=c(1,2,3),cutoff_label = c("Low","High"))

# Percentile NRI 
#perc_nri_train<-pct.nri_XL(risks.1 = rs_train$diab_rs,risks.2 = rs_train$TEU_rs,outcome = rs_train$TEU_T2DM_status,probs = c(0,0.5,0.9,1),confint = FALSE
                          # ) # too slow if want CI

# Below produces the same point estimate as pct.nri_XL but much faster for CI computation!
set.seed(100)
perc_nri_train<-nribin(event=rs_train$T2D10_status,
        p.std = rs_train$diab_rs_binary,p.new = rs_train$Eth_rs_binary,updown = 'category',
        cut = 2)


perc_nri_train_tb<-as.data.frame(perc_nri_train$nri)[1:3,]%>%
  mutate(NRI=c("Overall NRI","NRI for cases","NRI for noncases"),
         Estimate=pretty_dp(Estimate,3),
         `95% CI`=pretty_confint(Lower,Upper,dp=3))%>%
  select(NRI,Estimate,`95% CI`)
  
rownames(perc_nri_train_tb)=NULL  

```

```{r}
pander(perc_nri_train_tb,caption = "NRI results comparing the Leicester score and the enhanced score (without PRS) computed using the training data") 

```



**Comparison 2: Enhanced scores (without PRS) vs enhanced scores (with PRS)**:

We performed the similar steps:

* Step 1: Categorise individuals based on the enhanced scores (without PRS) to two risk groups: "Low:< `r Eth_threshold`" and "High: $\ge$ `r Eth_threshold`".

* Step 2: Categorise individuals based on the enhanced scores (with PRS) to two risk groups: "Low:< `r PRS_threshold`" and "High: $\ge$ `r PRS_threshold`".

* Step 3: Compute the categorical NRI with the incident T2D within 10 years. 


```{r NRI training 2,echo=FALSE}

# Categorise individuals to low (1) and high risk (2) groups

# reclassification table
reclassification_table(rs_train, match("T2D10_status",names(rs_train)), rs_train$Eth_rs_binary, rs_train$PRS_rs_binary, cutoff=c(1,2,3),cutoff_label = c("Low","High"))

# Percentile NRI 
#perc_nri_train<-pct.nri_XL(risks.1 = rs_train$diab_rs,risks.2 = rs_train$TEU_rs,outcome = rs_train$TEU_T2DM_status,probs = c(0,0.5,0.9,1),confint = FALSE
                          # ) # too slow if want CI

# Below produces the same point estimate as pct.nri_XL but much faster for CI computation!
set.seed(100)
perc_nri_train2<-nribin(event=rs_train$T2D10_status,
        p.std = rs_train$Eth_rs_binary,p.new = rs_train$PRS_rs_binary,updown = 'category',
        cut = 2)

perc_nri_train2_tb<-as.data.frame(perc_nri_train2$nri)[1:3,]%>%
  mutate(NRI=c("Overall NRI","NRI for cases","NRI for noncases"),
         Estimate=pretty_dp(Estimate,3),
         `95% CI`=pretty_confint(Lower,Upper,dp=3))%>%
  select(NRI,Estimate,`95% CI`)
  
rownames(perc_nri_train2_tb)=NULL  
```

```{r}
pander(perc_nri_train2_tb,caption = "NRI results comparing the enhanced score (without PRS) and the enhanced score (with PRS) computed using the training data")  

```

**Comparison 3: Leicester score vs enhanced scores (with PRS)**:

We performed the similar steps:

* Step 1: Categorise individuals based on the Leicester score to two risk groups: "Low:< 16" and "High: $\ge$ 16".

* Step 2: Categorise individuals based on the enhanced scores (with PRS) to two risk groups: "Low:< `r PRS_threshold`" and "High: $\ge$ `r PRS_threshold`".

* Step 3: Compute the categorical NRI with the incident T2D within 10 years. 


```{r NRI training 3,echo=FALSE}

# Categorise individuals to low (1) and high risk (2) groups

# reclassification table
reclassification_table(rs_train, match("T2D10_status",names(rs_train)), rs_train$diab_rs_binary, rs_train$PRS_rs_binary, cutoff=c(1,2,3),cutoff_label = c("Low","High"))

# Percentile NRI 
#perc_nri_train<-pct.nri_XL(risks.1 = rs_train$diab_rs,risks.2 = rs_train$TEU_rs,outcome = rs_train$TEU_T2DM_status,probs = c(0,0.5,0.9,1),confint = FALSE
                          # ) # too slow if want CI

# Below produces the same point estimate as pct.nri_XL but much faster for CI computation!
perc_nri_train3<-nribin(event=rs_train$T2D10_status,
        p.std = rs_train$diab_rs_binary,p.new = rs_train$PRS_rs_binary,updown = 'category',
        cut = 2)

perc_nri_train3_tb<-as.data.frame(perc_nri_train3$nri)[1:3,]%>%
  mutate(NRI=c("Overall NRI","NRI for cases","NRI for noncases"),
         Estimate=pretty_dp(Estimate,3),
         `95% CI`=pretty_confint(Lower,Upper,dp=3))%>%
  select(NRI,Estimate,`95% CI`)
  
rownames(perc_nri_train3_tb)=NULL  
```

```{r}
pander(perc_nri_train3_tb,caption = "NRI results comparing the Leicester score and the enhanced score (with PRS) computed using the training data")  

```



## Model performance in the test data


### AUC

We assess the outcome of T2D diagnosed within 10 years. Title: ROC Curve for risk scores using test data N= `r nrow(tot_test)`.

```{r AUC test CI}

AUC_test<-sapply(c("diab_rs","Eth_rs","PRS_rs"),function(i) AUC_CI(df=rs_test,score=i,status = "T2D10_status"))


AUC_test_label=c(
  paste0("LRA Score: ",AUC_test[1]),
  paste0("LRArev Score: ",AUC_test[2]),
  paste0("LRAprs Score: ",AUC_test[3])
)


# Try different format presenting ROC curve from below
ggplot(long_test,aes(d=T2D10_status, m=Value,color=Score))+geom_roc(n.cuts = 0,linealpha = 1,size=1)+
  style_roc()+
  #scale_color_discrete()+
  scale_color_manual(name="AUC (95% CI)",labels =  AUC_test_label,values = c("#F8766D","#00BA38", "#6193FF")) +
  labs(title ="",caption = "")

```


### NRI

**Comparison 1: Leicester scores vs enhanced scores (without PRS)**:

* Step 1: Categorise individuals based on the Leicester risk scores to two risk groups: "Low: <16" and "High: $\ge$ 16". Note: 16 is the established risk threshold for the Leicester risk score.

* Step 2: Categorise individuals based on enhanced scores (without PRS) to two risk groups: "Low:< `r Eth_threshold`" and "High: $\ge$ `r Eth_threshold`". Note: the cutoff is derived using the training data.

* Step 4: Compute the categorical NRI with the incident T2D within 10 years. 


```{r NRI test,echo=FALSE}


# reclassification table
reclassification_table(rs_test, match("T2D10_status",names(rs_test)), rs_test$diab_rs_binary, rs_test$Eth_rs_binary, cutoff=c(1,2,3),cutoff_label = c("Low","High"))

# Percentile NRI 
#perc_nri_train<-pct.nri_XL(risks.1 = rs_train$diab_rs,risks.2 = rs_train$TEU_rs,outcome = rs_train$TEU_T2DM_status,probs = c(0,0.5,0.9,1),confint = FALSE
                          # ) # too slow if want CI

# Below produces the same point estimate as pct.nri_XL but much faster for CI computation!
set.seed(100)
perc_nri_test<-nribin(event=rs_test$T2D10_status,
        p.std = rs_test$diab_rs_binary,p.new = rs_test$Eth_rs_binary,updown = 'category',
        cut = 2)

perc_nri_test_tb<-as.data.frame(perc_nri_test$nri)[1:3,]%>%
  mutate(NRI=c("Overall NRI","NRI for cases","NRI for noncases"),
         Estimate=pretty_dp(Estimate,3),
         `95% CI`=pretty_confint(Lower,Upper,dp=3))%>%
  select(NRI,Estimate,`95% CI`)
  
rownames(perc_nri_test_tb)=NULL  
```

```{r}
pander(perc_nri_test_tb,caption = "NRI results comparing the Leicester score and the enhanced score (without PRS) computed using the test data")  

```


**Comparison 2: Enhanced score (without PRS) vs enhanced score (with PRS)**

```{r NRI test 2,echo=FALSE}

# Categorise individuals to low (1) and high risk (2) groups

# reclassification table
reclassification_table(rs_test, match("T2D10_status",names(rs_test)), rs_test$Eth_rs_binary, rs_test$PRS_rs_binary, cutoff=c(1,2,3),cutoff_label = c("Low","High"))

# Percentile NRI 
#perc_nri_train<-pct.nri_XL(risks.1 = rs_train$diab_rs,risks.2 = rs_train$TEU_rs,outcome = rs_train$TEU_T2DM_status,probs = c(0,0.5,0.9,1),confint = FALSE
                          # ) # too slow if want CI

# Below produces the same point estimate as pct.nri_XL but much faster for CI computation!
set.seed(100)
perc_nri_test2<-nribin(event=rs_test$T2D10_status,
        p.std = rs_test$Eth_rs_binary,p.new = rs_test$PRS_rs_binary,updown = 'category',
        cut = 2)

perc_nri_test2_tb<-as.data.frame(perc_nri_test2$nri)[1:3,]%>%
  mutate(NRI=c("Overall NRI","NRI for cases","NRI for noncases"),
         Estimate=pretty_dp(Estimate,3),
         `95% CI`=pretty_confint(Lower,Upper,dp=3))%>%
  select(NRI,Estimate,`95% CI`)
  
rownames(perc_nri_test2_tb)=NULL  
```

```{r}
pander(perc_nri_test2_tb,caption = "NRI results comparing the enhanced score (without PRS) and the enhanced score (with PRS) computed using the test data")  

```

**Comparison 3: Leicester score vs enhanced score (with PRS)**

```{r NRI test 3,echo=FALSE}


# reclassification table
reclassification_table(rs_test, match("T2D10_status",names(rs_test)), rs_test$diab_rs_binary, rs_test$PRS_rs_binary, cutoff=c(1,2,3),cutoff_label = c("Low","High"))

# Percentile NRI 
#perc_nri_train<-pct.nri_XL(risks.1 = rs_train$diab_rs,risks.2 = rs_train$TEU_rs,outcome = rs_train$TEU_T2DM_status,probs = c(0,0.5,0.9,1),confint = FALSE
                          # ) # too slow if want CI

# Below produces the same point estimate as pct.nri_XL but much faster for CI computation!
set.seed(100)
perc_nri_test3<-nribin(event=rs_test$T2D10_status,
        p.std = rs_test$diab_rs_binary,p.new = rs_test$PRS_rs_binary,updown = 'category',
        cut = 2)

perc_nri_test3_tb<-as.data.frame(perc_nri_test3$nri)[1:3,]%>%
  mutate(NRI=c("Overall NRI","NRI for cases","NRI for noncases"),
         Estimate=pretty_dp(Estimate,3),
         `95% CI`=pretty_confint(Lower,Upper,dp=3))%>%
  select(NRI,Estimate,`95% CI`)
  
rownames(perc_nri_test3_tb)=NULL  
```

```{r}
pander(perc_nri_test3_tb,caption = "NRI results comparing the Leicester score and the enhanced score (with PRS) computed using the test data")  

```





